### 2019-11-15 1-Preamble/4-Acknowledgement/acknowledgement.tex
Mise à jour de la dédicace et des remerciements. Quelle déception et que de tristesse…
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

During my PhD, I was lucky to teach regularly in Mathematics and Computer Science.
I warmly thank David Pichardie for his trust and support, and also François Schwarzentruber, Nathalie Bertrand and David Cachera at ENS de Rennes, and Romaric Gaudel and his colleagues at ENSAI. Thanks to them, I loved to teach these $3 \times 64$ hours.
% since October $2016$.
% I am trully honored to join \emph{the Département d'Informatique} at ENS de Rennes in next September $2019$.
Thanks to my amazing students, I am proud of you and of our interesting exchanges.
%  and I feel lucky to be able to keep working alongside some of you.
I am thankful to Amélie Stainer and Gilbert Gabillard at Lycée Chateaubriand and Joliot-Curie for useful discussions.
\\
\indent
I am grateful to amazing professors who passed on their passion for mathematics and computer science to me.
A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers in Marseille.
Then at ENS de Cachan ($2011$-$16$), I especially think about Frédéric P., Alain T., Claudine P., Florian de V. and Nicolas V. in Maths, to Hubert C.-L., Jean G.-L., Sylvain S., Paul G., Serge H. in Computer Science, and to Carine S.-P., Delphine L. and Nicolas P.\textsuperscript{$\dagger$} for their support.
Thanks to Didier C. and Vijaysekhar C. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$) for our collaborations.

I would also like to thank some unusual things.
I thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday, despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} that gave me smiles and energy, and my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket), \LaTeX, Bash, Julia and Python, GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
Some people thank (a/some) God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work. I am trying hard to do better from now on!

>>>>

During my PhD, I was lucky to teach regularly in Mathematics and Computer Science.
I warmly thank David Pichardie for his trust and support, and also François Schwarzentruber, Nathalie Bertrand and David Cachera at ENS de Rennes, and Romaric Gaudel and his colleagues at ENSAI. Thanks to them, I loved to teach these $3 \times 64$ hours.
% since October $2016$.
% I am trully honored to join \emph{the Département d'Informatique} at ENS de Rennes in next September $2019$.
Thanks to my amazing students, I am proud of you and of our interesting exchanges.
%  and I feel lucky to be able to keep working alongside some of you.
I am thankful to Amélie Stainer and Gilbert Gabillard at Lycée Chateaubriand and Joliot-Curie for useful discussions.
\\
\indent
I am grateful to amazing professors who passed on their passion for mathematics and computer science to me,
% A partial list 
including Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers in Marseille.
Then at ENS de Cachan ($2011$-$16$), I especially think about Frédéric P., Alain T., Claudine P., Florian de V. and Nicolas V. in Maths, to Hubert C.-L., Jean G.-L., Sylvain S., Paul G., Serge H. in Computer Science, and to Carine S.-P., Delphine L. and Nicolas P.\textsuperscript{$\dagger$} for their support.
Thanks to Didier C. and Vijaysekhar C. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$) for our collaborations.

I would also like to thank some unusual things.
I thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday, despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} that gave me smiles and energy, and my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git, GitHub and Bitbucket, \LaTeX, Bash, Julia and Python, GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
Some people thank (a/some) God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work. I am trying hard to do better from now on!

### 2019-11-15 2-Chapters/1-Chapter/chapter1.tex
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

\item
    \emph{Decentralized Spectrum Learning for Radio Collision Mitigation in Ultra-Dense IoT Networks: LoRaWAN Case Study and Measurements},
    by Christophe Moy, \textbf{Lilian Besson}, Guillaume Delbarre \& Laurent Toutain,
    July $2019$.
    See Chapter~\ref{chapter:4}.
    Submitted for a special volume of \href{https://annalsoftelecommunications.wp.imt.fr/}{the Annals of Telecommunications} journal, on ``Machine Learning for Intelligent Wireless Communications and Networking''.
    % Preprint at \href{https://HAL.Inria.fr/hal-XXX}{\texttt{HAL.Inria.fr/hal-XXX}}.
    \cite{MoyBesson2019Annales}

@@

\item
    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
    by \textbf{Lilian Besson} \& Émilie Kaufmann,
    February $2019$.\\
    See Chapter~\ref{chapter:6}.
    Preprint at \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
    \cite{Besson2019GLRT}

    \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}

\item
    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
    by \textbf{Lilian Besson}, active development since October $2016$,
    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
    It currently consists in about $45000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
    and a complete documentation accessible on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} or \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.
    See Chapter~\ref{chapter:3}.
    \cite{SMPyBanditsJMLR,SMPyBandits}

>>>>

\item
    \emph{Decentralized Spectrum Learning for Radio Collision Mitigation in Ultra-Dense IoT Networks: LoRaWAN Case Study and Measurements},
    by Christophe Moy, \textbf{Lilian Besson}, Guillaume Delbarre \& Laurent Toutain,
    July $2019$.
    See Chapter~\ref{chapter:4}.
    % Preprint at \href{https://HAL.Inria.fr/hal-XXX}{\texttt{HAL.Inria.fr/hal-XXX}}.
    \cite{MoyBesson2019Annales}\\
    Submitted for a special volume of \href{https://annalsoftelecommunications.wp.imt.fr/}{the Annals of Telecommunications} journal, on ``Machine Learning for Intelligent Wireless Communications and Networking''.

\item
    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
    by \textbf{Lilian Besson} and others, active development since October $2016$,
    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
    It currently consists in about $45000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
    and a complete documentation accessible on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} or \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.\\
    See Chapter~\ref{chapter:3}.
    \cite{SMPyBanditsJMLR,SMPyBandits}\\
    Submitted for a special track of the \href{http://jmlr.org/mloss}{Journal of Machine Learning Research}, for Machine Learning Open-Source Software (MLOSS), in October $2019$.

\item
    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
    by \textbf{Lilian Besson} \& Émilie Kaufmann,
    February $2019$.\\
    See Chapter~\ref{chapter:6}.
    Preprint at \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
    \cite{Besson2019GLRT}\\
    An updated version, written with Odalric-Ambrym Maillard, was submitted to \href{http://www.aistats.org}{AISTATS 2020} in October $2020$.

@@

### 2019-11-15 2-Chapters/5-Chapter/chapter5.tex
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

One can observe in Table~\ref{table:5:meanRegretSimulationsEstimatingM} the empirical performances of different algorithms in an example problem, in each case of low, medium and maximum number of players ($M=3,6,9$ for $K=9$ arms).
The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar increase on their regret when they have to estimate $M$ on the fly (written ``Estimate $M$'' in red, in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.
For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large,
% (on the three algorithms and two cases, about $\times 3.4$),
and consistent on the different algorithms.
The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proven to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
In this extreme case, \RhoRand{} and its extension both perform very closely, as expected, and much worse than \MCTopM.
%
Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension of \MCTopM-\klUCB.

@@

>>>>

One can observe in Table~\ref{table:5:meanRegretSimulationsEstimatingM} the empirical performances of different algorithms in an example problem, in each case of low, medium and maximum number of players ($M=3,6,9$ for $K=9$ arms).
The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar increase on their regret when they have to estimate $M$ on the fly (written ``Estimate $M$'' in red, in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.
For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large,
% (on the three algorithms and two cases, about $\times 3.4$),
and consistent on the different algorithms.\newline
The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proven to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
In this extreme case, \RhoRand{} and its extension both perform very closely, as expected, and much worse than \MCTopM.
%
Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension of \MCTopM-\klUCB.

@@

\vfill{}

\hr{}
\begin{small}
    \begin{quote}
        % https://fr.wikiquote.org/wiki/Kaamelott/
        \emph{-- Victoriae mundis et mundis lacrima.\\
            Bon, ça ne veut absolument rien dire, mais je trouve que c’est assez dans le ton.}\\
        Le Roi Loth, interprété par François Rollin,\\
        \emph{Kaamelott}, Livre IV, ``Le désordre et la nuit''.

        \vspace*{10pt}

        \emph{-- C'est pas moi qu'explique mal, c'est les autres qui sont cons !}\\
        Perceval, interprété par Franck Pitiot,\\
        \emph{Kaamelott}, Livre IV, ``Perceval Fait Raitournelle''.
    \end{quote}
\end{small}

### 2019-11-15 2-Chapters/6-Chapter/chapter6.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006471
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

\input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex}

@@

\input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB_appendix.tex}

>>>>

\input{2-Chapters/6-Chapter/nonstatbandits/NonStatB.tex}

@@

\input{2-Chapters/6-Chapter/nonstatbandits/NonStatB_appendix.tex}

\vfill{}

\hr{}
\begin{small}
    \begin{quote}
        % https://fr.wikiquote.org/wiki/Kaamelott/
        \emph{-- Par exemple, vous prenez aujourd’hui. Vous comptez sept jours. Ça vous emmène dans une semaine. Et bien on sera exactement le même jour qu’aujourd’hui\dots{} À une vache près, hein… C’est pas une science exacte.}\\
        Karadoc, interprété par Jean-Christophe Hembert,\\
        \emph{Kaamelott}, Livre II, ``Sept cent quarante-quatre''.

        % \emph{-- Les chiffres, c'est pas une science exacte figurez-vous !}\\
        % Karadoc, interprété par Jean-Christophe Hembert,\\
        % \emph{Kaamelott}, Livre V, ``Le dernier recours''.
    \end{quote}
\end{small}
